Malacology: A Programmable Storage System

### 2016-10-14 paper/paper.tex
paper: tidy up malacology section, fix figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics{figures/overview.png}

\caption{Scalable storage systems consist of object storage daemons (OSD) which
store data fragments, monitor daemons (MON) that maintain cluster state, and
service-specific daemons such as file system metadata servers (MDS). Malacology
enables the programmability of internal abstractions and services (indicated by
bold arrows) in order to leverage the back-end system's durability, consistent
versioning, consensus, and metadata subsystems.  With Malacology, we add two
new services, ZLog and Mantle, that sit alongside the traditional user-facing
APIs (file, block, and object) in the storage stack.  \label{fig:overview}}
\end{figure}

@@

{\it \textbf{The definition a programmable storage system}} to be "a storage
system that allows exisiting abstractions to be re-used and composed to form
new data and metadata management services whose operation is orothogonal to
system correctness".  Our prototype includes (but is not limited to) the
following building blocks:

@@

\item An interface for managing strongly-consistent time-varying service
metadata.

\item An interface for installing and evolving domain-specific, cluster-wide data
interfaces.

\item An interface for managing access to shared resources using a variety of
optimization strategies.

\item An interface for load balancing resources across the cluster.

\item A durability interface for persisting policies using the underyling storage
stack's object store

@@

\section{Application-Specific Storage Stacks
}\label{highly-tailored-and-application-specifc-storage-systems}

@@

The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to use programmability because their needs, whether they be
related to performance, availability, consistency, convenience, etc., are not
satisfied by the existing default set of interfaces. The popularity of the custom object
interface facility of Ceph could be due to a number of reasons, such as the
default algorithms/tunables of the storage system are insufficient for the
application's performance goals, programmers want to exploit
application-specific semantics, and/or programmers know how to manage resources
to improve performance. Their solution is a way to work around the
traditionally rigid storage APIs because custom object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

@@

\begin{table*}
\centering
\title{Hi}
\begin{tabular}{  l | l | l | l }
\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Versioning}
  & Zookeeper coordination~\cite{hunt_zookeeper_2010}, Chubby~\cite{website:chubby}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols~\cite{CITEME}
  & serialization \& batching
  \\
{\it Logical Device}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
Beacon
  & Apache RabbitMQ~\cite{CITEME}
  & inter-daemon messages, centralized logging
  & debugging \& monitoring
  \\
\end{tabular}
\caption{The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces contributed in this paper.}
\label{table:examples}
\end{table*}

@@

\subsection{Versioning: Cluster State Management}
\label{sec:mon}
\label{consistencyversioning-of-cluster-state}

% straw man example
Keeping track of the state of a distributed system is an essential part of any 
successful service and a necessary component in order to diagnose and detect 
failures, when they occur. In the case of Ceph, a consistent view of cluster 
state among server daemons and clients is critical to provide strong consistency 
guarantees to clients. Ceph maintains cluster state information in per-subsystem 
``maps'' (e.g. OSD map, MDS map) that record membership and status information. 
A PAXOS-based monitoring service (termed, the Monitor) is responsible for 
integrating state changes into cluster maps, responding to requests from 
out-of-date clients and synchronizing members of the cluster whenever there is a 
change in a map so that they all observe the same system state. As a fundamental 
building block of many system designs, consensus abstractions such as PAXOS are 
a common technique for maintaining consistent data versions, and are a useful 
system to expose.

The default behavior of the monitor can be seen as a PAXOS-based notification 
system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing 
clients to identify when new values (termed epochs in Ceph) are associated to 
given maps. While Ceph does not expose this service directly, it does allow 
limited use by clients through a key-value service intended for low-volume, 
small size read-only configuration information. While a key-value service is 
generally useful, it doesn't provide many of the useful services hidden within 
the monitoring framework that would be required for applications with more 
demanding requirements, such as creating arbitrary maps or associating 
application-specific logic that can be executed for particular maps (or values 
in a map).

\paragraph*{Malacology:} Malacology exposes Ceph's PAXOS-based monitoring
service implementing a generic API
for adding arbitrary values to existing subsystem cluster maps, namely to the 
OSD map and the MDS map. As a consequence of this, applications can define 
simple but useful domain-specific logic to the PAXOS state machine, such as 
authorization control (just specific clients can write new values) or to trigger 
actions based on specific values (e.g. sanitize values). Since the monitor is 
intended to be out of high-performance I/O paths, a general guideline is to make 
use of this functionality infrequently and to assign small values to maps. 
Malacology itself makes use of this functionality to register, version and 
propagate dynamic code (Lua scripts) for new object interfaces defined in OSDs 
(\S\ref{active-storage}) and load balancing policies in MDSs 
(\S\ref{malacology:mds}).

\subsection{Logical Storage Devices}
\label{active-storage}

Briefly described in Section~\ref{sec:intro}, Ceph supports the creation
of application-specific object interfaces~\cite{weil_rados_2007}. The ability
to offload computation can reduce data movement, and transnational interfaces can
significantly simplify construction of complex storage interfaces that require
uncoordinated parallel access.

An object interface is a plugin structured in a similar way to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is invoked. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output. A wide range of native interfaces are available to
developers, such as reading and writing to a byte stream, controlling snapshots
and clones, all of which may be transactionally composed with access to a
sorted key-value service allowing applications to create rich interfaces, such
as using a read-modify-write guard to protect a shared index defined over
content store in a byte stream.

\begin{figure}[htbp]
\centering
\includegraphics{figures/implementation.png}
\caption{Malacology re-uses Ceph subsystems to allow users to dynamically define 
new object interfaces and load balancing policies. It achieves this by extending 
the OSD and MDS subsystems with an embedded Lua VM. To manage newly dynamic 
interface classes, it exposes the core functionality of the Monitor to allow 
arbitrary data to be associated to ODS and MDS cluster maps. After new 
interfaces are submitted to the monitor and the associated Paxos proposal is 
accepted by the monitor quorum, the script is redundantly and consistently 
distributed to OSDs so this snippets persist.
\label{fig:implementation}}
\end{figure}

\paragraph*{Malacology:} The implementation of Ceph's object abstraction,
although powerful, has some drawbacks. Supporting only C/C++ for object
interface developers, Ceph requires distribution of compiled binaries for the
correct architecture, adding a large barrier of entry for developers and
system administrators. Second, having no way to dynamically unload modules,
any changes require a full restart of an OSD which may have serious
performance impacts due to a loss of cached data. And
finally, the security limitations of the framework limit the use of object
classes to all but those with administrative level access and deep technical
expertise.

To address these concerns, Malacology takes advantage of community contributed
Lua extensions to Ceph allowing new object interfaces to be dynamically loaded
into the system and modified at runtime, resulting in a object storage API
with economy of expression, which at the same time provides the full set of
features of the base object class. New object interfaces that are
expressed in thousands of lines of code can be implemented in approximately an
order of magnitude less code~\cite{geambasu_comet_2010}.

{\bf Interface management.} All of these modifications work in tandem to
implement the desired behavior, and are typically co-designed together such
that each depends on the other to behave as expected. In highly dynamic
environments such as a distributed storage system, changes to the system
propagate at variable rates, requiring all components of the system to be able
to adapt. Thus, system services realized through extensibility should be
versioned together as a group, exposing the version to allow introspection and
domain-specific treatment of version mismatch scenarios.

And finally, it is important to not underestimate the importance of the
preservation of the modifications that enable a new system service. In many
(if not all) cases, interfaces defining access to data are just as important
as the data itself by virtue of inherently providing structural context. Thus,
all components of a system service must be kept to the same standards of
protection as data in the system itself.

To this end we have extended the monitoring service using the framework
described in Section~\ref{sec:mon} to provide an interface for registration
and verification of dynamically defined object interfaces. Using this service
guarantees that interface definitions are not only made durable, but are
transparently and consistently propagated throughput the cluster and that
clients are properly synchronized with the latest interfaces.

\subsection{Distributed Metadata Services}
\label{malacology:mds}

The distributed metadata service in Ceph provides clients with a POSIX file
system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
file systems protect resources by providing hierarchical indexing and
distributed locking services.  In Ceph, the locking service implements a
capability-based system that express what data and metadata clients are allowed
to access as well as what state they may cache and modify locally. While
designed for a fixed file abstraction, indexing, locking, and caching are all
common services that are useful to a broad spectrum of applications.

Ceph also addresses the challenge of balancing metadata load using load
balancing policies that compute metrics based on system state (e.g.  CPU and
memory utilization) and statistics collected by the cluster (e.g. the
popularity of an inode). Ceph uses dynamic subtree partitioning to move
variable sized namespace subtrees. These units can be shipped anywhere (i.e.,
to any MDS of any capacity) at any time for any reason. The original balancer
was designed with hard-coded policies and tunables.  Distributed applications
that share centralized resources (e.g. a database or directory) face similar
challenges which are often solved using application-specific sharding.

@@

\paragraph*{Malacology:} New interface hooks are added in strategic spots in
the metadata service to control load balancing and capabilities, and for
defining new file types allowing the indexing service to be re-purposed for
other types of named resources.

{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
metadata load balancer, has been implemented on Malacology so that its
infrastructure can enjoy all the properties of existing internal abstractions. The load
balancing logic is intact but the infrastructure has been
improved to safely load, manage, and persist its load balancing policies.

{\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
not enforced by the runtime, the system allows strings to be injected through
the admin daemon, constructing the Lua script in C++ is clunky, and the
administrator can inject really bad policies (e.g., while 1) that brings the
whole system down. Using the Malacology durability and beacon interfaces,
Mantle can safely load policies from the object store and centralize
warnings/errors, respectively.

{\bf Policy management.} A new interface is added to the metadata service
subsystem using the monitoring framework to support management of interfaces.
This ensures that the correct balancer version is consistently distributed and
verified on all MDS nodes. Policies are also durably saved with Ceph's
replication and redundancy.

@@

{\bf Capabilities.} The metadata service also manages sessions with its
clients. This gives both subsystems (MDSs and clients) the ability to cache
data and grab locks. The metadata service is responsible for responding to
client requests for capabilities by first revoking caps from other clients.
Currently, clients voluntarily release their capabilities, and the MDS maintains
a queue of requests.  Both clients and MDSs use a best-effort approach to
revoking or releasing capabilities.

{\bf File Types.} We leverage the same active and typed storage module
presented in DataMods~\cite{watkins_datamods_2012} to assign state and a set of
functions to a file (stored in the inode). This is valuable for
metadata load balancing so that certain types of files can have a higher
priority when making balancing decisions.

@@

\subsection{Durability}
\label{sec:durability}

>>>>

\begin{figure}[tb]
\centering
\includegraphics{figures/overview.png}
\caption{Scalable storage systems consist of storage daemons which store data
fragments, monitor daemons that maintain cluster state, and service-specific
daemons such as file system metadata servers. Malacology enables the
programmability of internal abstractions (indicated by bold arrows) in order to
re-use and compose existing subsystems.  With Malacology, we add two new
services, ZLog and Mantle, that sit alongside the traditional user-facing APIs
(file, block, and object) in the storage stack.  \label{fig:overview}}
\end{figure}

@@

{\it \textbf{The definition a programmable storage system}} to be ``a storage
system that allows existing abstractions to be re-used and composed to form
new data and metadata management services whose operation is orothogonal to
system correctness".  Our prototype includes (but is not limited to) the
following building blocks:

@@

\item An interface for managing strongly-consistent time-varying **service
metadata**.

\item An interface for installing and evolving domain-specific, cluster-wide **data
interfaces**.

\item An interface for managing access to **shared resources** using a variety of
optimization strategies.

\item An interface for **load balancing** resources across the cluster.

\item An interface for **durability** that persists policies using the underyling storage
stack's object store.

@@

\section{Application-Specific Storage Stacks}
\label{sec:application-specifc-storage-stacks}

@@

The takeaway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to use programmability because their needs, whether they be
related to performance, availability, consistency, convenience, etc., are not
satisfied by the existing default set of interfaces. The popularity of the custom object
interface facility of Ceph could be due to a number of reasons, such as the
default algorithms/tunables of the storage system are insufficient for the
application's performance goals, programmers want to exploit
application-specific semantics, and/or programmers know how to manage resources
to improve performance. Their solution is a way to work around the
traditionally rigid storage APIs because custom object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

@@

\begin{table*}
\centering
\begin{tabular}{  l | l | l | l }
\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Service Metadata}
  & Zookeeper coordination~\cite{hunt_zookeeper_2010}, Chubby~\cite{website:chubby}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols~\cite{CITEME}
  & serialization \& batching
  \\
{\it Object \& Data}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
Beacon
  & Apache RabbitMQ~\cite{CITEME}
  & inter-daemon messages, centralized logging
  & debugging \& monitoring
  \\
\end{tabular}
\caption{The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces contributed in this paper.}
\label{table:examples}
\end{table*}

@@

\subsection{Service Metadata Interface}
\label{sec:mon}
\label{service-metadata}

% straw man example
Keeping track of the state of a distributed system is an essential part of any
successful service and a necessary component in order to diagnose and detect
failures, when they occur. This is further complicated by variable propagation
delays and heterogenous hardware in highly dynamic environments.

%Thus, system
%services realized through extensibility should be versioned together as a
%group, exposing the version to allow introspection and domain-specific
%treatment of version mismatch scenarios.

In the case of Ceph, a consistent view of cluster state among server daemons
and clients is critical to provide strong consistency guarantees to clients.
Ceph maintains cluster state information in per-subsystem ``maps'' that record
membership and status information.  A PAXOS-based monitoring service is
responsible for integrating state changes into cluster maps, responding to
requests from out-of-date clients and synchronizing members of the cluster
whenever there is a change in a map so that they all observe the same system
state. As a fundamental building block of many system designs, consensus
abstractions such as PAXOS are a common technique for maintaining consistent
data versions, and are a useful system to expose.

The default behavior of the monitor can be seen as a PAXOS-based notification
system, similar to the one introduced in~\cite{burrows_chubby_2006}, allowing
clients to identify when new values (termed epochs in Ceph) are associated to
given maps. While Ceph does not expose this service directly, it does allow
limited use by clients through a key-value service intended for low-volume,
small size read-only configuration information. While a key-value service is
generally useful, it doesn't provide many of the useful services hidden within
the monitoring framework that would be required for applications with more
demanding requirements, such as creating arbitrary maps or associating
application-specific logic that can be executed for particular maps (or values
in a map).  Since the monitor is intended to be out of high-performance I/O
paths, a general guideline is to make use of this functionality infrequently
and to assign small values to maps.

\paragraph*{Malacology:} Malacology exposes a strongly-consistent view of
time-varying service metadata as a service rather than a hidden internal
component. Malacology provides a generic API for adding arbitrary values to
existing subsystem cluster maps. As a consequence of this, applications can
define simple but useful service-specific logic to the strongly-consistent
interface, such as authorization control (just specific clients can write new
values) or to trigger actions based on specific values (e.g. sanitize values).
The higher level services we implement in \S\ref{sec:services} make use of this
functionality to register, version and propagate dynamic code (Lua scripts) for
new object interfaces defined in storage daemons (\S\ref{active-storage}) and
policies in the load balancer \S\ref{malacology:mds}).  Using this service
guarantees that interface definitions are not only made durable, but are
transparently and consistently propagated throughput the cluster and that
clients are properly synchronized with the latest interfaces.

%And finally, it is important to not underestimate the importance of the
%preservation of the modifications that enable a new system service. In many
%(if not all) cases, interfaces defining access to data are just as important
%as the data itself by virtue of inherently providing structural context. Thus,
%all components of a system service must be kept to the same standards of
%protection as data in the system itself.

\subsection{Object \& Data Interface}
\label{object-data-interface}

Briefly described in Section~\ref{sec:application-specifc-storage-stacks}, Ceph
supports the creation of application-specific object
interfaces~\cite{weil_rados_2007}. The ability to offload computation can
reduce data movement, and transactional interfaces can significantly simplify
construction of complex storage interfaces that require uncoordinated parallel
access.

An object interface is a plugin structured in a similar way to that of an RPC
in which a developer creates a named function within the cluster that clients
may invoke. In the case of Ceph each function is implicitly run within the
context of an object specified when the function is called. Developers of
object interfaces express behavior by creating a composition of native
interfaces or other custom object interfaces, and handle serialization of
function input and output. A wide range of native interfaces are available to
developers, such as reading and writing to a byte stream, and controlling
snapshots and clones. These object interfaces may be transactionally composed
with access to a sorted key-value service allowing applications to create rich
semantics. An example is using a read-modify-write guard to protect a shared
index defined over content stored in a byte stream.

\begin{figure}[htbp]
\centering
\includegraphics{figures/implementation.png}
\caption{Malacology re-uses Ceph subsystems to allow users to dynamically
define new object/data interfaces and load balancing policies. It achieves this
by extending the object storage daemon (OSD) and metadata server daemon (MDS)
subsystems with an embedded Lua VM.  It uses the of the strongly-consistent
service metadata interface to propogate the object interfaces and versions
across the cluster.
\label{fig:implementation}} \end{figure}

\paragraph*{Malacology:} The implementation of Ceph's object abstraction,
although powerful, does not readily support programmability. Supporting only
C/C++ for object interface developers, Ceph requires distribution of compiled
binaries for the correct architecture, adding a large barrier of entry for
developers and system administrators. Second, having no way to dynamically
unload modules, any changes require a full restart of a storage daemon which
may have serious performance impacts due to loss of cached data. And finally,
the security limitations of the framework limit the use of object interfaces to
all but those with administrative level access and deep technical expertise.

To address these concerns, Malacology takes advantage of Lua extensions
contributed by the Ceph community. This allows new object interfaces to be
dynamically loaded into the system and modified at runtime, resulting in a
object storage API with economy of expression, which at the same time provides
the full set of features of the base object class. New object interfaces that
are expressed in thousands of lines of code can be implemented in approximately
an order of magnitude less code~\cite{geambasu_comet_2010}.

All of these modifications work in tandem to implement the desired behavior,
and are typically co-designed together such that each depends on the other to
behave as expected.

\subsection{Distributed Metadata Interfaces}
\label{malacology:mds}

The distributed metadata service in Ceph provides clients with a POSIX file
system abstraction~\cite{weil:sc2004-dyn-metadata}. In general, distributed
file systems protect resources by providing hierarchical indexing and
distributed locking services.  In Ceph, the locking service implements a
capability-based system that expresses what data and metadata clients are
allowed to access as well as what state they may cache and modify locally.
While designed for a fixed file abstraction, indexing, locking, and caching are
all common services that are useful to a broad spectrum of applications.
Distributed applications that share centralized resources (e.g. a database or
directory) face similar challenges which are often solved using
application-specific sharding.

Ceph also addresses the challenge of balancing metadata load with a separate
metadata cluster. This cluster uses load balancing policies to migrate
directory inodes around the cluster to alleviate load. The policies use metrics
based on system state (e.g.  CPU and memory utilization) and statistics
collected by the cluster (e.g. the popularity of an inode). Ceph uses dynamic
subtree partitioning to move variable sized namespace subtrees. These units can
be shipped anywhere (i.e., to any metadata server of any capacity) at any time
for any reason. The original balancer was designed with hard-coded policies and
tunables.

@@

\paragraph*{Malacology:} interfaces are added in strategic spots in the
metadata service for guarding shared resources and load balancing.

{\bf Shared Resource Interface.} The metadata service also manages sessions
with its clients. This gives both subsystems (metadata servers and clients) the
ability to cache data and obtain locks. The metadata service is responsible for
responding to client requests for capabilities by first revoking caps from
other clients.  Currently, clients voluntarily release their capabilities, and
the metadata server maintains a queue of requests.  Both clients and metadata
servers use a best-effort approach to revoking or releasing capabilities.

{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
metadata load balancer, has been implemented on Malacology so that its
infrastructure can enjoy all the properties of existing internal abstractions.
The load balancing logic is intact but the infrastructure has been improved to
safely load, manage, and persist its load balancing policies.  Composing Mantle
using the service metadata, durability and beacon interfaces, Mantle can safely
version balancer policies, save balancer policies in the back end object store
and centralize warnings/errors.

%{\bf Policy management.} A new interface is added to the metadata service
%subsystem using the monitoring framework to support management of interfaces.
%This ensures that the correct balancer version is consistently distributed and
%verified on all MDS nodes. Policies are also durably saved with Ceph's
%replication and redundancy.

@@

%{\bf File Types.} We leverage the same active and typed storage module
%presented in DataMods~\cite{watkins_datamods_2012} to assign state and a set of
%functions to a file (stored in the inode). This is valuable for
%metadata load balancing so that certain types of files can have a higher
%priority when making balancing decisions.

@@

\subsection{Durability Interface}
\label{sec:durability}
